---
title: "Graph Visualization on Pancreas"
author: "LylaAtta"
date: "9/3/2020"
output: html_document
---

Here I want to apply the projected neighbors graph visualization to the pancreas dataset that is used in the scVelo demo and compare it to the visualization on the U2OS dataset.  

```{r setup, include=FALSE}
#using velocyto2 evironment 
knitr::opts_chunk$set(echo = FALSE,include = TRUE)
library(reticulate)
library(igraph)
library(matie)
library(RANN)
library(velocyto.R)
library(MUDAN)
library(RSpectra)
source("projectedNeighbors.R")
```

Use the reticulate package to use scVelo from within R:   
```{r importing scvelo}
use_condaenv("velo",required = TRUE)
scv = import("scvelo")
scv$logging$print_version()
```
```{r loading panc data, message=FALSE}
adata = scv$datasets$pancreas()
#scv$pl$scatter(adata,legend_loc = "lower left", size = 60)


emb_umap = adata$obsm['X_umap'] #extract umap embedding 
clusters = adata$obs$clusters #extract clusters 
rownames(emb_umap) = names(clusters) = adata$obs_names$values

col = rainbow(length(levels(clusters)),s = 0.8, v = 0.8) 
cell.cols = col[clusters] #color according to cluster
names(cell.cols) = names(clusters)
```
```{r plot panc data on umap}
plot(emb_umap, col = cell.cols, pch=16, xlab = "UMAP X", ylab = "UMAP Y")
legend(x=13, y=0, legend = levels(clusters), col = col, pch = 16)
```

#### Compute velocities on pancreas data using velocyto
Extract spliced and unspliced data  
```{r extract counts, message = FALSE, results='hide'}
spliced = as.matrix(t(adata$layers['spliced']))
unspliced = as.matrix(t(adata$layers['unspliced']))
cells = adata$obs_names$values
genes = adata$var_names$values
colnames(spliced) = colnames(unspliced) = cells
rownames(spliced) = rownames(unspliced) = genes
```

Extract PCA coordinates  
```{r scvelo pca, message = FALSE, results='hide'}
pcs = adata$obsm['X_pca']
rownames(pcs) = cells
cell.dist = as.dist(1-cor(t(pcs))) #cell distance in PC space 
```

Filter genes  
```{r filter genes, message = FALSE }
gexpS = log10(rowSums(spliced)+1)
gexpU = log10(rowSums(unspliced)+2)
#plot(gexpS,gexpU)
goodGenes = genes[gexpS > 2 & gexpU > 1]
spliced = spliced[goodGenes,]
unspliced = unspliced[goodGenes,]
```

Downsample cells to make things easier  
```{r downsample, results='hide'}
spliced.sub = spliced[,seq(1,ncol(spliced),by=3)]
unspliced.sub = unspliced[,seq(1,ncol(unspliced),by=3)]
cell.dist.sub = as.dist(1-cor(t(pcs[seq(1,ncol(unspliced),by=3),])))
cell.cols.sub = cell.cols[colnames(spliced.sub)]
```

Normalize for dimensional reduction  
```{r normalize, results='hide'}
allExp = spliced.sub + unspliced.sub #use combined spliced and unspliced counts 
all.cpm = MUDAN::normalizeCounts(allExp) #cpm normalize
all.norm = MUDAN::normalizeVariance(all.cpm,details = TRUE, plot = TRUE) #variance stabilize 
odsGenes = goodGenes[all.norm$ods] #overdispersed genes
all.logODS = log10(as.matrix(t(all.norm$mat[all.norm$ods,]))+1) #keep overdispersed genes and log normalize


```

Dimensional reduction  
```{r dimensional reduction}
pca = svds(A = all.logODS, k=50, opts = list(center = TRUE, scale = FALSE, maxitr = 2000, tol = 1e-10))
var = pca$d
plot(var) # ~5 components 

emb.pca = pca$u[,1:2]
row.names(emb.pca) = row.names(all.logODS)
plot(emb.pca, col = cell.cols.sub, pch = 16, xlab = "PC1", ylab = "PC2")

```

Run velocyto on panc data  
```{r velocity, message = FALSE, results = 'hide'}
#vel = gene.relative.velocity.estimates(spliced.sub,unspliced.sub,deltaT=1,kCells=30,cell.dist=cell.dist.sub,fit.quantile=0.1,mult=100)
#saveRDS(vel,file = "panc_vel.rds")
vel = readRDS("panc_vel.rds")
curr = vel$current
proj = vel$projected

#show velocity on PCA
show.velocity.on.embedding.cor(scale(emb.pca), vel, n=100, scale='sqrt', cell.colors=cell.cols.sub,cex=1, arrow.scale=1, show.grid.flow=TRUE,min.grid.cell.mass=0.5, grid.n=30, arrow.lwd=2, main = "Velocities on PCA embedding")

#show velocity on UMAP 
emb.umap.sub = emb_umap[rownames(emb.pca),]
show.velocity.on.embedding.cor(scale(emb.umap.sub), vel, n=100, scale='sqrt', cell.colors=cell.cols.sub,cex=1, arrow.scale=1, show.grid.flow=TRUE,min.grid.cell.mass=0.5, grid.n=30, arrow.lwd=2, main = "Velocities on UMAP embedding")

```

#### Graph visualization 

Scores of observed and projected states in PC space  
```{r pc scores}
u = pca$u #scores
v = pca$v #loads
row.names(v) = odsGenes

common.genes = sapply(row.names(curr), function(x) x %in% odsGenes)
curr.pca = curr[common.genes,]
v.sub = v[row.names(curr.pca),]

par(mfrow = c(1,2))

curr.scores = t(curr.pca) %*% v.sub
plot(curr.scores[,1:2], col = cell.cols.sub, main = 'Observed',xlim = c(-0.55,0.25),ylim = c(-0.35,0.45))

proj.pca = proj[common.genes,]
proj.scores = t(proj.pca) %*% v.sub
plot(proj.scores[,1:2], col = cell.cols.sub, main = 'Projected',xlim = c(-0.55,0.25),ylim = c(-0.35,0.45))

```

Graph visualization on subset of cells from PC coordinates  
```{r graph viz from PCs}
set.seed(1)
cell.names = colnames(curr)
cell.names.sub = cell.names[seq(1,length(cell.names),5)] #downsample cells for speed 
ncells.sub = length(cell.names.sub)
cell.cols.grph = cell.cols[cell.names.sub]

curr.scores.cellsub = t(curr.scores[cell.names.sub,]) #### change here to change number of PCs included 
proj.scores.cellsub = t(proj.scores[cell.names.sub,]) ####

graphViz(curr.scores.cellsub,proj.scores.cellsub,10,"L2","cosine",0.25,cell.cols.grph)
```

Graph visualization on subset of cells from gene expression  
using `common.genes` (intersect of oveerdispersed genes, `odsGenes`, and genes in velocity output)  
```{r graph viz from gene exp}
set.seed(1)
curr.sub = curr[common.genes,cell.names.sub]
proj.sub = proj[common.genes,cell.names.sub]

graphViz(curr.sub,proj.sub,10,"L2","cosine",0.25,cell.cols.grph)


```

#### Graph parameters 
Effects of changing k, distance measure, similarity measure, and similarity threshold:  
Using PC generated graph  
  
L1 vs L2 as distance measure:  
```{r changing distance measure, echo = TRUE}
#using k=10, similarity=cosine, threshold=0.25
set.seed(1)
graphViz(curr.scores.cellsub,proj.scores.cellsub,10,"L1","cosine",0.25,cell.cols.grph,"L1 distance")
graphViz(curr.scores.cellsub,proj.scores.cellsub,10,"L2","cosine",0.25,cell.cols.grph,"L2 distance")
```

Pearson correlation vs Cosine similarity:  
```{r changing similarity measure, echo = TRUE}
#using k=10, distance=L2, threshold=0.25
set.seed(1)
graphViz(curr.scores.cellsub,proj.scores.cellsub,10,"L2","cosine",0.25,cell.cols.grph,"Cosine Similarity")
#graphViz(curr.scores.cellsub,proj.scores.cellsub,10,"L2","pearson",0.25,cell.cols.grph,"Pearson Correlation") #need to debug 
```

Number of out edges k:  
```{r changing k, echo = TRUE}
#using distance=L2, similarity=cosine, threshold=0.25
ks = c(1,5,10,20,50)
for (k in ks){
  graphViz(curr.scores.cellsub,proj.scores.cellsub,k,"L2","cosine",0.25,cell.cols.grph,paste("K = ",k))
}

```

Similarity threshold: 
```{r changing similarity threshold, echo = TRUE}
#using k=10, distance=L2, similarity=cosine
ts = c(-1, 0, 0.25, 0.5)
for (t in ts){
  graphViz(curr.scores.cellsub,proj.scores.cellsub,10,"L2","cosine",t,cell.cols.grph,paste("Similarity Threshold = ",t))
}

```



