---
title: "Visualizing cell cycle trajectory in MERFISH data using VeloViz"
author: "Lyla Atta, Jean Fan"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Visualizing cell cycle trajectory in MERFISH data using VeloViz}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Visualizing MERFISH data using VeloViz
In this vignette, we will use VeloViz to create a 2D embedding to visualize MERFISH data collected from U-2 OS cells in culture. Since this data comes from a cell line in culture, we expect the main temporal signal to be progression throught the cell-cycle. We will compare the VeloViz embedding to simply using the first two principal components. We will also compare the results we get when we restrict the input genes. The data used for this example is available with the VeloViz package and was initially obtained from the Xia et. al., *PNAS*, 2019.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide')
library(veloviz)
```

## Load preprocessed data  
MERFISH data from Xia et. al., *PNAS*, 2019. This data is provided with the VeloViz package.  
```{r load data}
col = MERFISH$col
pcs = MERFISH$pcs
vel = MERFISH$vel

curr = vel$current
proj = vel$projected
```

## Load cell cycle genes  
We will compare the VeloViz embeddings constructed with all genes to embeddings created using cell-cycle genes in the GO mitotic cell-cycle gene set as well as to embeddings created using genes that were found to have cell-cycle dependent expression in Xia et. al., *PNAS*, 2019.  
```{r cell cycle genes}
merfish.genes = rownames(curr)

#GO cell cycle genes (GO:0000278)
# https://www.gsea-msigdb.org/gsea/msigdb/cards/GO_MITOTIC_CELL_CYCLE
cycle.genes.go = read.csv("GO_0000278.csv",header = FALSE)
cycle.genes.go = cycle.genes.go$V1

merfish.cycle.go =  merfish.genes[which(merfish.genes %in% cycle.genes.go)]
curr.go = curr[merfish.cycle.go,]
proj.go = proj[merfish.cycle.go,]

#MERFISH genes exhibiting cell-cycle-dependent expression (Xia et al 2019, Supp Dataset 8)
# https://www.pnas.org/content/116/39/19490
cycle.genes.pnas = read.csv("pnas_sd08.csv",header = TRUE)
cycle.genes.pnas = cycle.genes.pnas$Gene

merfish.cycle.pnas = merfish.genes[which(merfish.genes %in% cycle.genes.pnas)]
curr.pnas = curr[merfish.cycle.pnas,]
proj.pnas = proj[merfish.cycle.pnas,]

```


# Build VeloViz embedding using all genes
```{r all genes, fig.height=7}
veloviz.all = buildVeloviz(
  curr = curr,
  proj = proj,
  normalize.depth = TRUE,
  use.ods.genes = FALSE,
  pca = TRUE,
  nPCs = 3,
  center = TRUE,
  scale = TRUE,
  k = 100,
  similarity.threshold = 0,
  distance.weight = 1,
  distance.threshold = 1,
  weighted = TRUE,
  seed = 0,
  verbose = FALSE
)

emb.all.vv = veloviz.all$fdg_coords

par(mfrow = c(1,2))
plotEmbedding(pcs[,1:2], colors = col[rownames(emb.all.vv)], main = 'all genes - PCA')
plotEmbedding(emb.all.vv, colors = col[rownames(emb.all.vv)], main = 'all genes - veloviz')

```


# Build VeloViz embedding using GO cell cycle genes
First, reduce dimensions.   
```{r pca go cell cycle genes, eval = FALSE}
curr.go.norm = normalizeDepth(curr.go)
curr.go.norm = normalizeVariance(curr.go.norm, details = TRUE)
curr.go.norm = log10(curr.go.norm$matnorm + 1)
curr.go.pca = RSpectra::svds(A = t(as.matrix(curr.go.norm)), k = 50,
                             opts = list(center = TRUE, scale = FALSE,
                                         maxitr = 2000, tol = 1e-10))
curr.go.pca = curr.go.pca$u
rownames(curr.go.pca) = rownames(pcs)

```

Now, build VeloViz embedding.
```{r go cell cycle veloviz, fig.height=7, eval = FALSE}
veloviz.go = buildVeloviz(
  curr = curr.go,
  proj = proj.go,
  normalize.depth = TRUE,
  use.ods.genes = FALSE,
  pca = TRUE,
  nPCs = 3,
  center = TRUE,
  scale = TRUE,
  k = 20,
  similarity.threshold = 0,
  distance.weight = 0.1,
  distance.threshold = 1,
  weighted = TRUE,
  seed = 0,
  verbose = FALSE
)

emb.go.vv = veloviz.go$fdg_coords
par(mfrow = c(1,2))
# plotEmbedding(curr.go.pca[,1:2], colors = col,
#               main = 'GO cell cycle genes - PCA')
plotEmbedding(emb.go.vv, colors = col[rownames(emb.go.vv)],
              main = 'GO cell cycle genes - veloviz')


```


# Build VeloViz embedding with cell-cycle dependent genes  
First, reduce dimensions.  
```{r pca pnas cell cycle genes, eval = FALSE}
curr.pnas.norm = normalizeDepth(curr.pnas)
curr.pnas.norm = normalizeVariance(curr.pnas.norm, details = TRUE)
curr.pnas.norm = log10(curr.pnas.norm$matnorm + 1)
curr.pnas.pca = RSpectra::svds(A = t(as.matrix(curr.pnas.norm)), k = 50,
                               opts = list(center = TRUE, scale = FALSE,
                                           maxitr = 2000, tol = 1e-10))
curr.pnas.pca = curr.pnas.pca$u
rownames(curr.pnas.pca) = rownames(pcs)

```

Now, build VeloViz embedding.  
```{r pnas cell cycle veloviz, fig.height=7, eval = FALSE}
veloviz.pnas = buildVeloviz(
  curr = curr.pnas,
  proj = proj.pnas,
  normalize.depth = TRUE,
  use.ods.genes = FALSE,
  pca = TRUE,
  nPCs = 3,
  center = TRUE,
  scale = TRUE,
  k = 50,
  similarity.threshold = 0.5,
  distance.weight = 0.01,
  distance.threshold = 1,
  weighted = TRUE,
  seed = 0,
  verbose = FALSE
)

emb.pnas.vv = veloviz.pnas$fdg_coords
par(mfrow = c(1,2))
# plotEmbedding(curr.pnas.pca[,1:2], colors = col,
#               main = 'Xia et al cell cycle genes - PCA')
plotEmbedding(emb.pnas.vv, colors = col[rownames(emb.pnas.vv)],
              main = 'Xia et al cell cycle genes - veloviz')
```


\newpage
# Comparing VeloViz embeddings
```{r compare veloviz, fig.width=10, eval = FALSE}
par(mfrow = c(1,3))

plotEmbedding(emb.all.vv, colors = col[rownames(emb.pnas.vv)],
              main = 'all genes - veloviz')

plotEmbedding(emb.go.vv, colors = col[rownames(emb.pnas.vv)],
              main = 'GO cell cycle genes - veloviz')

plotEmbedding(emb.pnas.vv, colors = col[rownames(emb.pnas.vv)],
              main = 'Xia et al cell cycle genes - veloviz')




```
