---
title: "Visualizing pancreas development using VeloViz"
author: "Lyla Atta, Jean Fan"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Visualizing pancreas development using VeloViz}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide', message = FALSE, warning = FALSE)
library(veloviz)
```

# Visualization using VeloViz  
In this vignette, we will create a velocity-informed 2D embedding of pancreas development scRNAseq using VeloViz. We will compare the embedding created from the full (subsampled) data to that created from data missing developmental intermediates to see if we can still see similar cellular trajectories. The data used for this example is available with the VeloViz package and was initially obtained from the scVelo package (Bergen et. al. *Nature Biotech*, 2020, Bastidas-Ponce et. al. *Development*, 2019). More details about the data, preprocessing, and normalization steps are available in the tutorials at jef.works/veloviz/pancreas. 

1) Load example full data from VeloViz  
```{r load from veloviz}
clusters = pancreas$clusters # cell type annotations
pcs = pancreas$pcs # PCs used to make other embeddings (UMAP,tSNE..)
vel = pancreas$vel # velocity

#choose colors based on clusters for plotting later
cell.cols = rainbow(8)[as.numeric(clusters)]
names(cell.cols) = names(clusters)

#extract current observed and predicted future transcriptional 
#states from velocity object
curr = vel$current
proj = vel$projected


```


2) Build VeloViz graph from current and projected using `buildVeloviz`  
To build the graph, we have to specify multiple parameters that control the features of the graph:  
**`k`**: how many out-edges each cell can have  
**`similarity_threshold`**: cosine similarity threshold specifying how similar the velocity and cell transition vectors have to be for an out-edge to be included  
**`distance_weight`**: weight for distance component of composite distance - with large weights, graph will prioritize linking cells where projected states and neighbors are close in PC space; with small weights, graph will prioritize linking cells where velocity and cell transition vectors are most similar  
**`distance_threshold`**: quantile threshold specifying minimum distance in PC space between projected state and neighbor for out-edge to be included - e.g. a distance threshold of 0.2 means that any edges where the distance component is not in the smallest 20% of distances in PC space will be removed from the graph   
**`weighted`**: whether to use composite distance to determine graph edge weights (`TRUE`) or to assign all edges equal weights (`FALSE`)    
```{r buildVeloviz}
veloviz = buildVeloviz(
  curr = curr, proj = proj,
  normalize.depth = TRUE,
  use.ods.genes = TRUE,
  alpha = 0.05,
  pca = TRUE,
  nPCs = 20,
  center = TRUE,
  scale = TRUE,
  k = 5,
  similarity.threshold = 0.25,
  distance.weight = 1,
  distance.threshold = 0.5,
  weighted = TRUE,
  seed = 0,
  verbose = FALSE
)
```


3) Plot VeloViz graph using a force-directed graph layout
```{r plot Veloviz}
emb.veloviz = veloviz$fdg_coords #2D coordinates
plotEmbedding(emb.veloviz, groups=clusters[rownames(emb.veloviz)], main='veloviz')

par(mfrow=c(1,1), mar=rep(1,4))
g = plotVeloviz(veloviz, clusters=clusters[rownames(emb.veloviz)], seed=0, verbose=TRUE)

```


# Visualization with missing intermediates using VeloViz  
Load data: this is the same dataset as above but missing a proportion of Ngn3 high EP cells  
```{r get data missing}
clusters = pancreasWithGap$clusters # cell type annotations
pcs = pancreasWithGap$pcs # PCs used to make other embeddings (UMAP,tSNE..)
vel = pancreasWithGap$vel # velocity

#choose colors based on clusters for plotting later
cell.cols = rainbow(8)[as.numeric(clusters)]
names(cell.cols) = names(clusters)
```

Create VeloViz embedding  
```{r veloviz missing}
curr = vel$current
proj = vel$projected

veloviz = buildVeloviz(
  curr = curr, proj = proj,
  normalize.depth = TRUE,
  use.ods.genes = TRUE,
  alpha = 0.05,
  pca = TRUE,
  nPCs = 20,
  center = TRUE,
  scale = TRUE,
  k = 5,
  similarity.threshold = 0.25,
  distance.weight = 1,
  distance.threshold = 0.5,
  weighted = TRUE,
  seed = 0,
  verbose = FALSE
)
```

Plot VeloViz embedding 
```{r plot missing}
emb.veloviz = veloviz$fdg_coords
plotEmbedding(emb.veloviz, groups=clusters[rownames(emb.veloviz)], main='veloviz')

par(mfrow=c(1,1), mar=rep(1,4))
g = plotVeloviz(veloviz, clusters=clusters[rownames(emb.veloviz)], seed=0, verbose=TRUE)
```



