---
title: "Pancreas example"
author: "Jean Fan"
date: "11/13/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide')
library(veloviz)
library(reticulate)
library(velocyto.R)
```

# Visualization using VeloViz  
## Preprocessing  
Inputs to VeloViz are the scores in PCA space of the current and projected transcriptional states, which we get here by calculating RNA velocity using velocyto.  
To get current and projected PC scores from raw counts, we first follow standard filtering, normalization, and dimensional reduction steps and then calculate velocity. (Steps 1-4 can be skipped by loading the example dataset in the VeloViz package - see 4*).    

0) Get Data:  
```{r get data}
#getting pancreas data from scVelo
use_condaenv("cellrank", required = TRUE)
scv = import("scvelo")
adata = scv$datasets$pancreas()

#extract spliced, unspliced counts
spliced <- as.matrix(Matrix::t(adata$layers['spliced']))
unspliced <- as.matrix(Matrix::t(adata$layers['unspliced']))
cells <- adata$obs_names$values
genes <- adata$var_names$values
colnames(spliced) <- colnames(unspliced) <- cells
rownames(spliced) <- rownames(unspliced) <- genes

#clusters
clusters <- adata$obs$clusters 
names(clusters) <- adata$obs_names$values

#subsample to make things faster 
set.seed(0)
good.cells <- sample(cells, length(cells)/5)
spliced <- spliced[,good.cells]
unspliced <- unspliced[,good.cells]
clusters <- clusters[good.cells]

dim(spliced)
dim(unspliced)

```

\newpage

1) Filter good genes  
```{r filter}
#keep genes with >10 total counts
good.genes = genes[rowSums(spliced) > 10 & rowSums(unspliced) > 10]
spliced = spliced[good.genes,]
unspliced = unspliced[good.genes,]

dim(spliced)
dim(unspliced)
```



2) Normalize  
```{r normalize}
counts = spliced + unspliced # use combined spliced and unspliced counts
cpm = normalizeDepth(counts) # normalize to counts per million 
varnorm = normalizeVariance(cpm) # variance stabilize, find overdispersed genes
lognorm = log10(varnorm + 1) # log normalize

```



3) Reduce Dimensions  
After filtering and normalizing, we reduce dimensions, and calculate cell-cell distance in PC space. This distance will be used to compute velocity.  
```{r pca dist}
#PCA on centered and scaled expression of overdispersed genes
pcs = reduceDimensions(lognorm, center = TRUE, scale = TRUE, nPCs = 50)

#cell distance in PC space
cell.dist = as.dist(1-cor(t(pcs))) # cell distance in PC space
```

\newpage
## Velocity  
4) Calculate velocity  
Next, we compute velocity from spliced and unspliced counts and cell-cell distances using velocyto. This will give us the current and projected transcriptional states.  
```{r velocity}
vel = gene.relative.velocity.estimates(spliced,
                                       unspliced,
                                       kCells = 30,
                                       cell.dist = cell.dist,
                                       fit.quantile = 0.1)
```

4*) Load example from VeloViz  
This example dataset is available with the veloviz package.  
```{r load from veloviz}
clusters = pancreas$clusters # cell type annotations
pcs = pancreas$pcs # PCs used to make other embeddings (UMAP,tSNE..)
vel = pancreas$vel # velocity
```


5) Normalize current and projected  
Now that we have the current and projected expression, we want to go through a similar normalization process as we did with the raw counts and then reduce dimensions in PCA. Steps 5-7 can be done together using the `buildVeloviz` function (see 7*).  
```{r normalize current and projected}
curr = vel$current
proj = vel$projected

#normalize depth
curr.norm = normalizeDepth(curr)
proj.norm = normalizeDepth(proj)

#variance stabilize current
curr.varnorm.info = normalizeVariance(curr.norm, details = TRUE) 
curr.varnorm = curr.varnorm.info$matnorm

#use same model for projected
scale.factor = curr.varnorm.info$df$scale_factor #gene scale factors 
names(scale.factor) = rownames(curr.varnorm.info$df)

m = proj.norm
rmean = Matrix::rowMeans(m) #row mean
sumx = Matrix::rowSums(m)
sumxx = Matrix::rowSums(m^2)
rsd = sqrt((sumxx - 2 * sumx * rmean + ncol(m) * rmean ^ 2) / (ncol(m)-1)) #row sd

proj.varnorm = proj.norm / rsd * scale.factor[names(rsd)]
proj.varnorm = proj.norm[rownames(curr.varnorm),]

```



6) Project current and projected into PC space  
```{r pca current and projected}
#log normalize 
curr.pca = log10(curr.varnorm + 1)
proj.pca = log10(proj.varnorm + 1)

#mean center
c.rmean = Matrix::rowMeans(curr.pca)
curr.pca = curr.pca - c.rmean
p.rmean = Matrix::rowMeans(proj.pca)
proj.pca = proj.pca - p.rmean

#scale variance
c.sumx = Matrix::rowSums(curr.pca)
c.sumxx = Matrix::rowSums(curr.pca^2)
c.rsd = sqrt((c.sumxx - 2*c.sumx*c.rmean + ncol(curr.pca)*c.rmean^2)/(ncol(curr.pca)-1))
curr.pca = curr.pca/c.rsd

p.sumx = Matrix::rowSums(proj.pca)
p.sumxx = Matrix::rowSums(proj.pca^2)
p.rsd = sqrt((p.sumxx - 2*p.sumx*p.rmean + ncol(proj.pca)*p.rmean^2)/(ncol(proj.pca)-1))
proj.pca = proj.pca/p.rsd

#PCA
pca = RSpectra::svds(A = Matrix::t(curr.pca), k=20,
                     opts = list(
                       center = FALSE, ## already done
                       scale = FALSE, ## already done
                       maxitr = 2000,
                       tol = 1e-10))
  
#scores of current and projected
curr.scores = Matrix::t(curr.pca) %*% pca$v[,1:10]
proj.scores = Matrix::t(proj.pca) %*% pca$v[,1:10]

```



\newpage
## VeloViz
7) Build graph using VeloViz  
Now we can use the PC projections of the current and projected transcriptional states to build the VeloViz graph. To build the graph, we have to specify multiple parameters that control the features of the graph:  
**`k`**: how many out-edges each cell can have  
**`similarity_threshold`**: cosine similarity threshold specifying how similar the velocity and cell transition vectors have to be for an out-edge to be included  
**`distance_weight`**: weight for distance in PC space between projected state and neighbor - with large weights, graph will prioritize linking cells where projected states and neighbors are close in PC space; with small weights, graph will prioritize linking cells where velocity and cell transition vectors are similar  
**`distance_threshold`**: quantile threshold specifying minimum distance in PC space between projected state and neighbor for out-edge to be included  
**`weighted`**: whether to use composite distance to determine graph edge weights (`TRUE`) or to assign all edges equal weights (`FALSE`).    


```{r veloviz}
#VeloViz graph parameters 
k = 5
similarity.threshold = 0.25
distance.weight = 1
distance.threshold = 0.5
weighted = TRUE

#build graph
set.seed(0)
veloviz = graphViz(t(curr.scores), t(proj.scores), k,
                   cell.colors=NA,
                   similarity_threshold=similarity.threshold,
                   distance_weight = distance.weight,
                   distance_threshold = distance.threshold,
                   weighted = weighted,
                   plot = FALSE,
                   return_graph = TRUE)


emb.veloviz = veloviz$fdg_coords
plotEmbedding(emb.veloviz, groups=clusters[rownames(emb.veloviz)], main='veloviz')

par(mfrow=c(1,1), mar=rep(1,4))
g = plotVeloviz(veloviz, clusters=clusters[rownames(emb.veloviz)], seed=0, verbose=TRUE)

```

7*) Build VeloViz graph from current and projected using `buildVeloviz`  
```{r buildVeloviz}
curr = vel$current
proj = vel$projected

veloviz = buildVeloviz(
  curr = curr, proj = proj,
  normalize.depth = TRUE, 
  use.ods.genes = TRUE,
  alpha = 0.05,
  pca = TRUE,
  nPCs = 20,
  center = TRUE,
  scale = TRUE, 
  k = 5,
  similarity.threshold = 0.25,
  distance.weight = 1,
  distance.threshold = 0.5,
  weighted = TRUE,
  seed = 0,
  verbose = FALSE
)

emb.veloviz = veloviz$fdg_coords
plotEmbedding(emb.veloviz, groups=clusters[rownames(emb.veloviz)], main='veloviz')

par(mfrow=c(1,1), mar=rep(1,4))
g = plotVeloviz(veloviz, clusters=clusters[rownames(emb.veloviz)], seed=0, verbose=TRUE)

```



\newpage 
## Compare to other embeddings
```{r, fig.height = 7, fig.width=8}
par(mfrow = c(2,2))

#PCA
emb.pca = pcs[,1:2]
plotEmbedding(emb.pca, groups=pancreas$clusters, main='PCA')

#tSNE
set.seed(0)
emb.tsne = Rtsne::Rtsne(pcs, perplexity=30)$Y
rownames(emb.tsne) = rownames(pcs)
plotEmbedding(emb.tsne, groups=pancreas$clusters, main='tSNE', 
              xlab = "t-SNE X", ylab = "t-SNE Y")

##UMAP
set.seed(0)
emb.umap = uwot::umap(pcs, min_dist = 0.5)
rownames(emb.umap) <- rownames(pcs)
plotEmbedding(emb.umap, groups=pancreas$clusters, main='UMAP',
              xlab = "UMAP X", ylab = "UMAP Y")

#veloviz
plotEmbedding(emb.veloviz, groups=clusters[rownames(emb.veloviz)], main='veloviz')


```

\newpage

# Visualization with missing intermediates using VeloViz  
Load data: this is the same dataset as above but missing a proportion of Ngn3 high EP cells  
```{r get data missing}
clusters = pancreasWithGap$clusters # cell type annotations
pcs = pancreasWithGap$pcs # PCs used to make other embeddings (UMAP,tSNE..)
vel = pancreasWithGap$vel # velocity
```

Create VeloViz embedding  
```{r veloviz missing}
curr = vel$current
proj = vel$projected

veloviz = buildVeloviz(
  curr = curr, proj = proj,
  normalize.depth = TRUE, 
  use.ods.genes = TRUE,
  alpha = 0.05,
  pca = TRUE,
  nPCs = 20,
  center = TRUE,
  scale = TRUE, 
  k = 5,
  similarity.threshold = 0.25,
  distance.weight = 1,
  distance.threshold = 0.5,
  weighted = TRUE,
  seed = 0,
  verbose = FALSE
)

emb.veloviz = veloviz$fdg_coords
plotEmbedding(emb.veloviz, groups=clusters[rownames(emb.veloviz)], main='veloviz')

par(mfrow=c(1,1), mar=rep(1,4))
g = plotVeloviz(veloviz, clusters=clusters[rownames(emb.veloviz)], seed=0, verbose=TRUE)
```

Compare to other embeddings  
```{r, fig.height = 7, fig.width=8}
par(mfrow = c(2,2))

#PCA
emb.pca = pcs[,1:2]
plotEmbedding(emb.pca, groups=pancreas$clusters, main='PCA')

#tSNE
set.seed(0)
emb.tsne = Rtsne::Rtsne(pcs, perplexity=30)$Y
rownames(emb.tsne) = rownames(pcs)
plotEmbedding(emb.tsne, groups=pancreas$clusters, main='tSNE', 
              xlab = "t-SNE X", ylab = "t-SNE Y")

##UMAP
set.seed(0)
emb.umap = uwot::umap(pcs, min_dist = 0.5)
rownames(emb.umap) <- rownames(pcs)
plotEmbedding(emb.umap, groups=pancreas$clusters, main='UMAP',
              xlab = "UMAP X", ylab = "UMAP Y")

#veloviz
plotEmbedding(emb.veloviz, groups=clusters[rownames(emb.veloviz)], main='veloviz')


```




\newpage 



```{r getdata, eval=FALSE, include = FALSE}
use_condaenv("cellrank", required = TRUE)
scv <- import("scvelo")
adata <- scv$datasets$pancreas()

## spliced and unspliced expression matrices
spliced <- as.matrix(t(adata$layers['spliced']))
unspliced <- as.matrix(t(adata$layers['unspliced']))
cells <- adata$obs_names$values
genes <- adata$var_names$values
colnames(spliced) <- colnames(unspliced) <- cells
rownames(spliced) <- rownames(unspliced) <- genes
## extract clusters
clusters <- adata$obs$clusters 
names(clusters) <- adata$obs_names$values
## old embedding
emb.original <- adata$obsm['X_umap'] #extract umap embedding
rownames(emb.original) <- names(clusters)
## plot
par(mfrow <- c(1,1))
plotEmbedding(emb.original, groups = clusters, 
              xlab = "UMAP X", ylab = "UMAP Y", 
              mark.clusters = TRUE)

## subsample to create smaller dataset
## that can be included with package
set.seed(0)
good.cells <- sample(rownames(emb.original), nrow(emb.original)/5)
spliced <- spliced[,good.cells]
unspliced <- unspliced[,good.cells]
clusters <- clusters[good.cells]
emb.original <- emb.original[good.cells,]
## plot
par(mfrow = c(1,1))
plotEmbedding(emb.original, groups = clusters, 
              xlab = "UMAP X", ylab = "UMAP Y", 
              mark.clusters=TRUE)

## filter to well detected genes
vi <- rowSums(spliced) > 10 & rowSums(unspliced) > 10
spliced <- spliced[vi,]
unspliced <- unspliced[vi,]

## analyze
counts <- spliced + unspliced # use combined spliced and unspliced counts
cpm <- normalizeDepth(counts) # cpm normalize
matnorm <- normalizeVariance(cpm)
matnorm <- log10(matnorm + 1)
pcs <- reduceDimensions(matnorm, center = TRUE, scale = TRUE, nPCs = 50)

## velocity model
library(velocyto.R)
cell.dist <- as.dist(1-cor(t(pcs))) # cell distance in PC space
vel <- gene.relative.velocity.estimates(spliced,
                                       unspliced,
                                       kCells = 30,
                                       cell.dist = cell.dist,
                                       fit.quantile = 0.1)

## save
pancreas <- list(
  spliced = spliced,
  unspliced = unspliced,
  clusters = clusters,
  pcs = pcs,
  cell.dist = cell.dist,
  vel = vel
)
usethis::use_data(pancreas, overwrite=TRUE)
```



```{r fulleval, eval = FALSE, include = FALSE}
#Compare visualizations
## load data
library(veloviz)
data("pancreas")

par(mfrow=c(2,2), mar=rep(1,4))

## 2D embedding by PCA
emb.pcs = pancreas$pcs[,1:2]
plotEmbedding(emb.pcs, groups=pancreas$clusters, main='PCA')

## 2D embedding by tSNE
set.seed(0)
emb.tsne = Rtsne::Rtsne(pancreas$pcs[,1:10], perplexity=30)$Y
rownames(emb.tsne) <- rownames(pancreas$pcs)
plotEmbedding(emb.tsne, groups=pancreas$clusters, main='tSNE')

## 2D embedding by UMAP
set.seed(0)
emb.umap = uwot::umap(pancreas$pcs[,1:10], min_dist = 0.5)
rownames(emb.umap) <- rownames(pancreas$pcs)
plotEmbedding(emb.umap, groups=pancreas$clusters, main='UMAP')

## 2D embedding by veloviz
vig = buildVeloviz(
  curr = pancreas$vel$curr,
  proj = pancreas$vel$proj,
  normalize.depth = FALSE,
  use.ods.genes = TRUE,
  alpha = 0.05,
  pca = TRUE,
  nPCs = 10,
  center = TRUE,
  scale = TRUE,
  k = 5,
  seed = 0,
  verbose = TRUE
)
emb.veloviz = vig$fdg_coords
plotEmbedding(emb.veloviz, groups=pancreas$clusters, main='veloviz')

par(mfrow=c(1,1), mar=rep(1,4))
g = plotVeloviz(vig, clusters=pancreas$clusters, seed=0, verbose=TRUE)
```



```{r removecells, eval=FALSE, include = FALSE}
#Now we remove cells
## remove EP cells along original trajectory
x = emb.original[,1]
vi = x > -5 & x < 0
good.cells = rownames(emb.original)[!vi]
plotEmbedding(emb.original[good.cells,], groups=clusters, 
              xlab = "UMAP X", ylab = "UMAP Y", mark.clusters=TRUE)
spliced = spliced[,good.cells]
unspliced = unspliced[,good.cells]
clusters = clusters[good.cells]

## analyze
counts <- spliced + unspliced # use combined spliced and unspliced counts
cpm <- normalizeDepth(counts) # cpm normalize
matnorm <- normalizeVariance(cpm)
matnorm <- log10(matnorm + 1)
pcs <- reduceDimensions(matnorm, center = TRUE, scale = TRUE, nPCs = 50)

## velocity model
library(velocyto.R)
cell.dist <- as.dist(1-cor(t(pcs))) # cell distance in PC space
vel <- gene.relative.velocity.estimates(spliced,
                                       unspliced,
                                       kCells = 30,
                                       cell.dist = cell.dist,
                                       fit.quantile = 0.1)

pancreasWithGap <- list(
  spliced = spliced,
  unspliced = unspliced,
  clusters = clusters,
  pcs = pcs,
  cell.dist = cell.dist,
  vel = vel
)
usethis::use_data(pancreasWithGap, overwrite=TRUE)
```


```{r subeval, eval = FALSE, include = FALSE}
#Compare
## load data
library(veloviz)
data("pancreasWithGap")

par(mfrow=c(2,2), mar=rep(1,4))

## 2D embedding by PCA
emb.pcs = pancreasWithGap$pcs[,1:2]
plotEmbedding(emb.pcs, groups=pancreasWithGap$clusters, main='PCA')

## 2D embedding by tSNE
set.seed(0)
emb.tsne = Rtsne::Rtsne(pancreasWithGap$pcs[,1:10], perplexity=30)$Y
rownames(emb.tsne) <- rownames(pancreasWithGap$pcs)
plotEmbedding(emb.tsne, groups=pancreasWithGap$clusters, main='tSNE')

## 2D embedding by UMAP
set.seed(0)
emb.umap = uwot::umap(pancreasWithGap$pcs[,1:10], min_dist = 0.5)
rownames(emb.umap) <- rownames(pancreasWithGap$pcs)
plotEmbedding(emb.umap, groups=pancreasWithGap$clusters, main='UMAP')

## 2D embedding by veloviz
vig = buildVeloviz(
  curr = pancreasWithGap$vel$curr,
  proj = pancreasWithGap$vel$proj,
  normalize.depth = FALSE,
  use.ods.genes = TRUE,
  alpha = 0.05,
  pca = TRUE,
  nPCs = 10,
  center = TRUE,
  scale = TRUE,
  k = 5,
  seed = 0,
  verbose = FALSE
)
emb.veloviz = vig$fdg_coords
plotEmbedding(emb.veloviz, groups=pancreasWithGap$clusters, main='veloviz')

par(mfrow=c(1,1), mar=rep(1,4))
g = plotVeloviz(vig, clusters = pancreasWithGap$clusters, seed = 0)
```

