---
title: "Graph Visualization on Dentate Gyrus Data"
author: "LylaAtta"
date: "9/3/2020"
output: html_document
---

Here I want to apply the projected neighbors graph visualization to the neural crest dataset that is used in the scVelo demo.  

#### Setup and get data from scVelo
```{r setup, include=FALSE}
#using velocyto2 evironment 
knitr::opts_chunk$set(echo = FALSE)
library(reticulate)
library(igraph)
library(matie)
library(RANN)
library(velocyto.R)
library(MUDAN)
library(RSpectra)
library(umap)
source("projectedNeighbors.R")
```

Use the reticulate package to use scVelo from within R:   
```{r importing scvelo}
use_condaenv("velo",required = TRUE)
scv = import("scvelo")
#scv$logging$print_version()
```

```{r loading panc data, message=FALSE}
adata = scv$datasets$dentategyrus()
#scv$pl$scatter(adata,legend_loc = "lower left", size = 60)


emb_umap = adata$obsm['X_umap'] #extract umap embedding 
clusters = adata$obs$clusters #extract clusters 
rownames(emb_umap) = names(clusters) = adata$obs_names$values

col = rainbow(length(levels(clusters)),s = 0.8, v = 0.8) 
cell.cols = col[clusters] #color according to cluster
names(cell.cols) = names(clusters)
```

```{r plot panc data on umap}
plot(emb_umap, col = cell.cols, pch=16, xlab = "UMAP X", ylab = "UMAP Y", xlim = c(-15,25))
legend(x=15, y=15, legend = levels(clusters), col = col, pch = 16)
```

#### Compute velocities on neuro data using velocyto
Extract spliced and unspliced data  
```{r extract counts, message = FALSE, results='hide'}
spliced = as.matrix(t(adata$layers['spliced']))
unspliced = as.matrix(t(adata$layers['unspliced']))
cells = adata$obs_names$values
genes = adata$var_names$values
colnames(spliced) = colnames(unspliced) = cells
rownames(spliced) = rownames(unspliced) = genes
```


Filter genes  
```{r filter genes, message = FALSE, fig.height = 3, fig.width = 6}
gexpS = log10(rowSums(spliced)+1)
gexpU = log10(rowSums(unspliced)+1)

par(mfrow = c(1,2))
hist(gexpS, breaks = 500, main = "log spliced counts")
abline(v = 1.75, col = 'red')
hist(gexpU, breaks = 500, main = "log unspliced counts")
abline(v = 1, col = "red")

#plot(gexpS,gexpU)
goodGenes = genes[gexpS > 1.75 & gexpU > 1]
spliced = spliced[goodGenes,]
unspliced = unspliced[goodGenes,]
```

Downsample cells to make things easier  
```{r downsample, results='hide'}
spliced.sub = spliced[,seq(1,ncol(spliced),by=3)]
cell.names.sub = colnames(spliced.sub)
unspliced.sub = unspliced[,seq(1,ncol(unspliced),by=3)]
#cell.dist.sub = as.dist(1-cor(t(pcs[seq(1,ncol(unspliced),by=3),])))
cell.cols.sub = cell.cols[colnames(spliced.sub)]
```

Normalize for dimensional reduction  
```{r normalize, results='hide'}
allExp = spliced.sub + unspliced.sub #use combined spliced and unspliced counts 
all.cpm = MUDAN::normalizeCounts(allExp) #cpm normalize
all.norm = MUDAN::normalizeVariance(all.cpm,details = TRUE, plot = FALSE) #variance stabilize 
odsGenes = goodGenes[all.norm$ods] #overdispersed genes
all.logODS = log10(as.matrix(t(all.norm$mat[all.norm$ods,]))+1) #keep overdispersed genes and log normalize


```

Dimensional reduction  
```{r dimensional reduction, fig.height = 4, fig.width=7}
pca = svds(A = all.logODS, k=50, opts = list(center = TRUE, scale = FALSE, maxitr = 2000, tol = 1e-10))
var = pca$d
plot(var) # ~6 components 

emb.pca = pca$u[,1:2]
row.names(emb.pca) = row.names(all.logODS)
plot(emb.pca, col = cell.cols.sub, pch = 16, xlab = "PC1", ylab = "PC2", main = "PCA on subsampled data", xlim = c(-0.05,0.2), ylim = c(-0.2,0.1))
legend(x=0.13, y=0.1, legend = levels(clusters), col = col, pch = 16, cex = 0.8)

#calculate cell-cell distance in pc space for velocity calculation 
cell.dist = as.dist(1-cor(t(pca$u))) #cell distance in PC space 

```

```{r umap}
pcsToKeep = pca$u[,1:20]
emb.umap = umap(pcsToKeep)$layout
row.names(emb.umap) = row.names(all.logODS)
plot(scale(emb.umap),col = cell.cols.sub, pch = 16, xlab = "UMAP X", ylab = "UMAP Y",main = "UMAP on subsampled data", xlim = c(-4,7))
legend(x=3.5, y=3.5, legend = levels(clusters), col = col, pch =16, cex = 0.8)
```



Run velocyto on neuro data  
```{r velocity, message = FALSE, results = 'hide'}
#vel = gene.relative.velocity.estimates(spliced.sub,unspliced.sub,deltaT=1,kCells=30,cell.dist=cell.dist,fit.quantile=0.1,mult=100)
#saveRDS(vel,file = "./saved_objects/neuro_vel.rds")
vel = readRDS("./saved_objects/neuro_vel.rds")
curr = vel$current
proj = vel$projected

#show velocity on PCA
show.velocity.on.embedding.cor(scale(emb.pca), vel, n=100, scale='sqrt', cell.colors=cell.cols.sub,cex=1, arrow.scale=1, show.grid.flow=TRUE,min.grid.cell.mass=0.5, grid.n=30, arrow.lwd=2, main = "Velocities on PCA embedding")

#show velocity on UMAP 
emb.umap.sub = emb_umap[rownames(emb.pca),]
show.velocity.on.embedding.cor(scale(emb.umap.sub), vel, n=100, scale='sqrt', cell.colors=cell.cols.sub,cex=1, arrow.scale=1, show.grid.flow=TRUE,min.grid.cell.mass=0.5, grid.n=30, arrow.lwd=2, main = "Velocities on UMAP embedding")

```

#### Graph visualization 

Scores of observed and projected states in PC space  
```{r pc scores, fig.width= 5, fig.height=8}
u = pca$u #scores
v = pca$v #loads
row.names(v) = odsGenes

common.genes = sapply(row.names(curr), function(x) x %in% odsGenes)
common.genes.names = rownames(curr)[common.genes]
curr.pca = curr[common.genes,]
v.sub = v[row.names(curr.pca),]

par(mfrow = c(2,1))

curr.scores = t(curr.pca) %*% v.sub
plot(curr.scores[,1:2], col = cell.cols.sub, pch = 16, main = 'Observed',xlim = c(-0.5,4),ylim = c(-3,1.25))
legend(x=2, y=1.2, legend = levels(clusters), col = col, pch =16, cex = 0.8)

proj.pca = proj[common.genes,]
proj.scores = t(proj.pca) %*% v.sub
plot(proj.scores[,1:2], col = cell.cols.sub, pch = 16, main = 'Projected',xlim = c(-0.5,4),ylim = c(-3,1.25))
legend(x=2, y=1.2, legend = levels(clusters), col = col, pch =16, cex = 0.8)

```

Graph visualization on subset of cells from PC coordinates 
```{r graph viz from PCs, results = "hide"}
source("projectedNeighbors.R")
set.seed(1)
cell.names = colnames(curr)
cell.names.sub = cell.names[sample(c(1:length(cell.names)),500)] #downsample cells for speed 
#cell.names.sub = cell.names #no downsampling 
ncells.sub = length(cell.names.sub)
cell.cols.grph = cell.cols[cell.names.sub]

curr.scores.cellsub = t(curr.scores[cell.names.sub,]) #### change here to change number of PCs included 
proj.scores.cellsub = t(proj.scores[cell.names.sub,]) ####
plot(t(curr.scores.cellsub[1:2,]), col = cell.cols.grph, pch = 16)

k = 10
t = 0
set.seed(1)
graphViz(curr.scores.cellsub,proj.scores.cellsub,k,"L2","cosine",t,weighted = FALSE, cell.colors = cell.cols.grph,title = paste("K =",k,"unweighted, thresh =",t))
set.seed(1)
graphViz(curr.scores.cellsub,proj.scores.cellsub,k,"L2","cosine",t,weighted = TRUE, cell.colors = cell.cols.grph,title = paste("K =",k,"weighted, thresh =",t))
```

Graph visualization on subset of cells from gene expression  
using `common.genes` (intersect of overdispersed genes, `odsGenes`, and genes in velocity output (genes with high correlation b/w spliced and unspliced))  
```{r graph viz from gene exp, results = "hide"}
set.seed(1)
curr.sub = curr[common.genes,cell.names.sub]
proj.sub = proj[common.genes,cell.names.sub]

k =10
t = 0
graphViz(curr.sub,proj.sub,k,"L2","cosine",t,weighted = FALSE, cell.colors = cell.cols.grph,title = paste("K =",k,"unweighted, thresh =",t))
graphViz(curr.sub,proj.sub,k,"L2","cosine",t,weighted = TRUE, cell.colors = cell.cols.grph,title = paste("K =",k,"weighted, thresh =",t))

```

#### Graph parameters 
Effects of changing k, distance measure, similarity measure, and similarity threshold:  
Using PC generated graph  
  
L1 vs L2 as distance measure:  
```{r changing distance measure, echo = TRUE, results = "hide"}
#using k=10, similarity=cosine, threshold=0.25
k = 10
t = 0
set.seed(1)
graphViz(curr.sub,proj.sub,k,"L1","cosine",t,weighted = TRUE, cell.colors = cell.cols.grph,title = paste("L1 distance,K =",k,"weighted, thresh =",t))
set.seed(1)
graphViz(curr.sub,proj.sub,k,"L2","cosine",t,weighted = TRUE, cell.colors = cell.cols.grph,title = paste("L2 distance,K =",k,"weighted, thresh =",t))
```

Pearson correlation vs Cosine similarity:  
```{r changing similarity measure, results = "hide", echo = TRUE}
k = 10
t = -1
set.seed(1)
graphViz(curr.sub,proj.sub,k,"L2","cosine",t,weighted = TRUE, cell.colors = cell.cols.grph,title = paste("Cosine similarity,K =",k,"\nweighted, thresh =",t))
set.seed(1)
graphViz(curr.sub,proj.sub,k,"L2","pearson",t,weighted = TRUE, cell.colors = cell.cols.grph,title = paste("Pearson correlation,K =",k,"\nweighted, thresh =",t))
```
..looks like correlation is more conservative than cosine similarity.

Number of out edges k:  
```{r changing k, results = "hide",fig.width= 6, fig.height=6}
#using distance=L2, similarity=cosine, threshold=0.25
ks = c(1,5,10,20,50)
t = 0
for (k in ks){
  print(paste("k=",k))
  set.seed(1)
  graphViz(curr.sub,proj.sub,k,"L1","cosine",t,weighted = TRUE, cell.colors = cell.cols.grph,title = paste("K =",k,"weighted, thresh =",t))
}

```

Similarity threshold: 
```{r changing similarity threshold}
#using k=10, distance=L2, similarity=cosine
ts = c(-1, 0, 0.25, 0.5, 0.7, 0.9)
k = 10
for (t in ts){
  print(t)
  set.seed(1)
  graphViz(curr.sub,proj.sub,k,"L1","cosine",t,weighted = TRUE, cell.colors = cell.cols.grph,title = paste("K =",k,"weighted, thresh =",t))
}

```

#### Graph parameters consistency scores
Number of out edges k:  
L2 distance, cosine similarity, similarity threshold = 0  
```{r, results = "hide"}
#using distance=L2, similarity=cosine, threshold=0
ks = c(1,5,10,20,50,100)
conScores_k = matrix(NA,nrow = ncells.sub,ncol=length(ks))
for (i in seq(1,length(ks))){
  k = ks[i]
  currG = graphViz(curr.scores.cellsub,proj.scores.cellsub,k,"L2","cosine",0,cell.cols.grph,paste("K = ",k),plot = FALSE,return_graph = TRUE)
  currCoords = currG$fdg_coords
  curr.con = consistency(currCoords, deltaExp,20,FALSE)
  conScores_k[,i] = curr.con
  # col.gradient = colorRampPalette(c('blue','red'))
  # cons.col = col.gradient(ncells.sub)[as.numeric(cut(curr.con,breaks = ncells.sub))]
  # plot(currCoords,col = cons.col,pch = 16)
}

par(mfrow = c(3,2))
hist(conScores_k[,1],col="red",breaks = 150,xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="K = 1")
hist(conScores_k[,2],col="blue",breaks = 150, xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="K = 5")
hist(conScores_k[,3],col="green",breaks = 150, xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="K = 10")
hist(conScores_k[,4],col="yellow",breaks = 150, xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="K = 20")
hist(conScores_k[,5],col="black",breaks = 150, xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="K = 50") #add = TRUE
hist(conScores_k[,6],col="cyan",breaks = 150, xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="K = 100") #add = TRUE

print("Mean consistency scores")
print(colMeans(conScores_k))
print("Median consistency scores")
print(sapply(c(1:6), function(x) median(conScores_k[,x])))
```

  
Similarity threshold:  
```{r, results = "hide"}
#using k=10, distance=L2, similarity=cosine
ts = c(-1, 0, 0.25, 0.5)
conScores_t = matrix(NA,nrow = ncells.sub,ncol=length(ts))
for (i in seq(1,length(ts))){
  t = ts[i]
  currG = graphViz(curr.scores.cellsub,proj.scores.cellsub,10,"L2","cosine",t,cell.cols.grph,paste("Similarity Threshold = ",t),plot = FALSE,return_graph = TRUE)
  currCoords = currG$fdg_coords
  curr.con = consistency(currCoords, deltaExp,20,FALSE)
  conScores_t[,i] = curr.con
  # col.gradient = colorRampPalette(c('blue','red'))
  # cons.col = col.gradient(ncells.sub)[as.numeric(cut(curr.con,breaks = ncells.sub))]
  # plot(currCoords,col = cons.col,pch = 16)
}

par(mfrow = c(3,2))
hist(conScores_t[,1],col="red",breaks = 125,xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="t = -1")
hist(conScores_t[,2],col="blue",breaks = 125, xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="t = 0")
hist(conScores_t[,3],col="green",breaks = 125, xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="t = 0.25")
hist(conScores_t[,4],col="yellow",breaks = 125, xlim=c(-0.5,1),ylim=c(0,9),xlab = "Cell Consistency Score",main="t = 0.5")#add = TRUE

print("Mean consistency scores")
print(colMeans(conScores_t))
print("Median consistency scores")
print(sapply(c(1:4), function(x) median(conScores_t[,x])))

```

#### Consistency of fdg compared to other embeddings
Consistency score in FDG compared to PCA and UMAP computed on same cell subset  
```{r}
#pca and umap on subsample 
#pca
all.logODS.sub = all.logODS[cell.names.sub,]
pca.sub = svds(A = all.logODS.sub,k = 50, opts = list(center = TRUE, scale = FALSE, maxitr = 2000, tol = 1e-10))
var.sub = pca.sub$d
#plot(var.sub) # 5 components
emb.pca.sub = pca.sub$u[,1:2]
row.names(emb.pca.sub) = row.names(all.logODS.sub)
plot(emb.pca.sub,col = cell.cols.grph, pch = 16, xlab = "PC1", ylab = "PC2", main = "PCA on cell subset")


#umap
set.seed(1)
pcs.sub = pca.sub$u
emb.umap.sub = umap(emb.pca.sub)$layout
row.names(emb.umap.sub) = row.names(all.logODS.sub)
plot(scale(emb.umap.sub),col = cell.cols.grph, pch = 16, xlab = "UMAP X", ylab = "UMAP Y", main = "UMAP on cell subset")

```


```{r, results="hide"}
set.seed(1)

deltaExp = vel$deltaE[,cell.names.sub]

#emb.pca.sub.grph = emb.pca[cell.names.sub,]
pcaCon = consistency(emb.pca.sub,deltaExp,20,FALSE)

#emb.umap.sub.grph = emb.umap.sub[cell.names.sub,]
umapCon = consistency(emb.umap.sub,deltaExp,20,FALSE)

fdgCoords = graphViz(curr.scores.cellsub,proj.scores.cellsub,20,"L2","cosine",0.25,cell.cols.sub,plot = FALSE,return_graph = TRUE)$fdg_coords
fdgCon = consistency(fdgCoords,deltaExp,20,FALSE)

```
```{r, fig.width = 5, fig.height = 7}
par(mfrow = c(3,1))
x = c(-0.2,1)
y = c(0,15)
xt = "Cell Consistency Score"
hist(pcaCon,breaks = 100, col = "red",xlim = x, ylim = y, xlab =  xt, main = "PCA Embedding Consistency")
hist(umapCon,breaks = 100, col = "green",xlim = x, ylim = y,xlab =  xt, main = "UMAP Embedding Consistency")
hist(fdgCon, breaks = 100, col = "blue",xlim = x, ylim = y,xlab =  xt, main = "FDG Embedding Consistency")

print("Mean consistency scores for PCA, UMAP, FDG")
mean(pcaCon)
mean(umapCon)
mean(fdgCon)
print("Median consistency scores for PCA, UMAP, FDG")
median(pcaCon)
median(umapCon)
median(fdgCon)

```

